Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-03-17 be.tex
Rewriting formula in appendix
Adding to appendix

Using equations~\eqref{eq:enConsInt},~\eqref{eq:massFlux},
and~\eqref{eq:Phi}, we may obtain an expression for the enthalpy,
$\ke+\gammaf \frac{p}{\rho}$ of the gas inside of the break radius
$\rb$.

\begin{align}
&\frac{v^2}{2}+\gammaf \frac{p}{\rho} = \frac{G \Mbh}{x \rs} + \frac{G
  M_\star}{x \rs} - U_o \frac{x^{2-\delta}-1}{2-\delta}\nonumber\\
&-U_o\left[\frac{1}{2-\delta}-\frac{x^{5-2\delta}-1}{x^{3-\delta}-1}\frac{3-\delta}{(5-2\delta)(2-\delta)}\right]\nonumber\\
&+\frac{v_w^2}{2}-\frac{G \Mbh}{2 r_s} \frac{3-\delta}{2-\delta}\frac{x^{2-\delta}-1}{x^{3-\delta}-1} -\frac{G M_\star}{2 r_s} \frac{x^{5-2\delta}-1}{x^{3-\delta}-1}\frac{3-\delta}{5-2\delta},
\end{align}

@@

\begin{multline}
  \frac{v^2}{2}+\gammaf \frac{p}{\rho}=\frac{G \Mbh}{\rs}
  \biggl[\zeta^2 w^{\frac{1}{3 -\delta}}+\frac{1}{x}+\frac{w}{x}
-\frac{1}{2} \frac{3-\delta}{2-\delta} \frac{x^{2 -\delta}-1}{x^{3-\delta}-1}\\
  -w (3 -\delta) \left(-\frac{3-\delta}{(5-2\delta)(2-\delta)}\frac{x^{5 -2 \delta}-1}{ \left(x^{3-\delta}-1\right)}\right)\\
  -\frac{w (3
    -\delta) x^{2 -\delta}}{2
    -\delta}-\frac{w (3 -\delta) \left(x^{5
        -2 \delta}-1\right)}{2 (5 -2 \delta)
    \left(x^{3 -\delta}-1\right)}
  \biggr]
\end{multline}
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "ms"
%%% End:

>>>>

Using equations~\eqref{eq:enConsInt},~\eqref{eq:massFlux},
and~\eqref{eq:Phi}, we may obtain an expression for the enthalpy,
$\ke+\gammaf \frac{p}{\rho}$ of the gas inside of the break radius
$\rb$. {\bf AG: 2s with question marks would not be included if the $G
\Mstar/r$ }

\begin{align}
&\frac{v^2}{2}+\gammaf \frac{p}{\rho} = \frac{G \Mbh}{x \rs} + \frac{G
  M_\star}{x \rs} - U_o \frac{x^{2-\delta}-1}{2-\delta}\nonumber\\
&-U_o\left[\frac{1}{2-\delta}-\frac{x^{5-2\delta}-1}{x^{3-\delta}-1}\frac{3-\delta}{(5-2\delta)(2-\delta)}\right]\nonumber\\
&+\frac{v_w^2}{2}-\frac{G \Mbh}{2 r_s} \frac{3-\delta}{2-\delta}\frac{x^{2-\delta}-1}{x^{3-\delta}-1} -\frac{G M_\star}{2? r_s} \frac{x^{5-2\delta}-1}{x^{3-\delta}-1}\frac{3-\delta}{5-2\delta},
\end{align}

@@

\begin{multline}
  \frac{v^2}{2}+\gammaf \frac{p}{\rho}
=\frac{G \Mbh}{\rs} 
\biggl[
  \zeta^2 w^{\frac{1}{3 -\delta}}+\frac{1}{x}+\frac{w}{x}
  -\frac{1}{2} \frac{3-\delta}{2-\delta} \frac{x^{2  -\delta}-1}{x^{3-\delta}-1}\\
  -w \frac{3 -\delta}{2 -\delta} x^{2 -\delta}
  -w \frac{(2-\delta)(3-\delta)-2?(3-\beta)^{2}}{(5-2\delta)(2-\delta)} \frac{x^{5-2\delta}-1}{x^{3-\delta}-1}
\biggr].
\label{eq:enthAnal}
\end{multline}

Recall that $\rs$ the stagnation radius, $x\equiv r/\rs$, $\zeta$ is
the ratio of the extra heating rate to the stellar velocity in of the
influence radius $\rsoi$, $w\equiv (\rs/\rsoi)^{3\delta}$, and
$\delta$ is the power law slope of the stellar density profile inside
of the break radius, $\rb$.

For a given $\zeta$ we define a critical $\rs/\rsoi=min{
  \frac{frac{7}{2}-\densSlope}{\densSlope \zeta^2, 1}}$ for thermal stability, 
where the first term in the brackets corresponds to twice the value
predicted by equation~\eqref{eq:rsRinf}, in the limit the stellar mass
as $\rs$ is small compared to $\Mbh$. This prescription is somehwat
ad hoc, but...{\bf AG--some motivation...}

Then from equation~\eqref{eq:enthAnal} we may obtain the critical
wind heating parameter $v_{w, \rm crit}/\sigma_{\rm inf}$ for thermal stability as
a function of $\rb/\rsoi$. This is shown in Fig.~\ref{fig:vwCrit}

\begin{figure}
\includegraphics[width=\columnwidth]{vwCrit.pdf}
\caption{\label{fig:vwCrit} Critical heating rate required for thermal
  stability normalized to the velocity dispersion at the sphere
  influence radius, $\sigma_{\rm soi}$ versus $\rb/\rsoi$, the ratio
  of the break radius in the stellar density profile to the sphere of
  influence radius. We assume a core stellar dnsity profile inside of
  $\rb$, so that $\rhostar\sim r^{-1-\Gamma}$ with $\Gamma=0.1$.}
\end{figure}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "ms"
%%% End:

### 2015-03-17 ms.tex
Adding to appendix

This paper is organized as followed.  In $\S\ref{sec:model}$ we
describe our model, including the sample of galaxies used in our
analysis ($\S\ref{sec:gal_model}$) and our numerical procedure for
calculating the steady-state hydrodynamic profile of the CNM
($\S\ref{sec:hydro}$).  In $\S\ref{sec:results}$ we describe some
analytic results, while in $\S\ref{sec:numerical}$ we describe our numerical results.  In $\S\ref{sec:discussion}$ we discuss the implications of our results, to topics including...   In $\S\ref{sec:conclusions}$ we summarize our conclusions.  Table \ref{table:definitions} provides the definitions of commonly used variables.  Appendices \ref{app:rs} and XXX provides useful analytic results, while Appendix \ref{app:windheat} describes our method for calculating stellar wind heating and mass input in greater detail.

@@

In the opposite limit of weak heating ($v_{w} \ll \sigma$), the
stagnation radius moves to large radii $\gg r_{\rm soi}$ or may not
exist at all, formally implying that the entire ISM of the galaxy is
accreting.  This result makes intuitive sense: gas is supplied to the
nuclear by stars which are gravitationally bound to the black hole, so
outflows are possible only if the specific heating rate $\sim
v_{w}^{2}$ significantly exceeds the specific gravitational binding
energy $\sim \sigma^{2}$.  In fact, for core galaxies the specific heating rate necessary to produce an unbound outflow can greatly exceed $\sigma^{2}$ due to the additional gravitational potential of stars outside the sphere of influence (Appendix XXX).

@@

Thermal stability ($\dot{q}_{\rm heat} > |\dot{q}_{\rm rad}|$; $t_{\rm
  cool} > t_{\rm ff}$) near the stagnation radius thus requires a
critical minimum heating rate
\begin{align}
v_{w} > v_{\rm TI} \simeq
  {\rm max}\begin{cases}
   200 \eta_{0.02}^{0.14} M_{\bullet,8}^{0.11}\,{\rm km\,s^{-1}}  &, \text{core}\\
   200 \eta_{0.02}^{0.17} M_{\bullet,8}^{0.08}\,{\rm km\,s^{-1}}   &, \text{cusp}\\
f(r_b/r_{\rm soi})\sigma,     
  \end{cases}
  \label{eq:cooling3}
\end{align}
The requirement $v_{\rm w} \gtrsim \sigma$ results because the top two lines are derived using equation (\ref{eq:rs_simple}) for $r_s$.  However, when $v_{\rm w} \lesssim f(r_b/r_{\rm soi})\sigma$ the stagnation radius moves to very large radii, or may not exist at all (see also Fig.~\ref{fig:stag}), resulting in much higher gas densities and inevitable cooling instability.  The dimensionless function f is approximately unity for cusp galaxies, but increases $\propto (r_b/r_{\rm soi})^{1/2}$ for core galaxies (Fig.~XXX).  
%Our numerical results in $\S\ref{sec:numerical}$ demonstrate that $\tilde{v}_{w} \gtrsim v_{\rm CI} \sim 200-300$ km s$^{-1}$ is generally sufficient for stability at all radii (Fig.~\ref{fig:cooling}).

@@

If such cyclic outbursts also manifest on the properties of gas on small, parsec scales, then the large scatter in the accretion rates and hence X-ray luminosities in Fig \ref{fig:miller} could result from the wide range of inflow rates experienced over the course of a cyclic episode between instabilities and periodic nuclear outbursts (e.g.~\citealt{Ciotti+10}).  Due to their deeper potential well, we find that core galaxies are more likely to possess higher accretion rates and potentially be thermally unstable for a fixed wind heating rate than cusp galaxies (Fig.~XXX).  This could help explain why core galaxies show a wide range of nuclear radio and X-ray luminosities (suggesting cyclic accretion episodes), while cusp galaxies are confined below a threshold (e.g., \citealt{Bender+89,   Pellegrini99, Capetti&Balmaverde05}), consistent with thermally-stable, steady-state picture.

@@

% \subsubsection{Comparison with observed X-ray Luminosity-Mass
%   Relation}
% %Note currently I reverse-engineer the stellar mass from the black
% %hole mass inferred from M-sigma; however, it would be better to use
% %the tabulated black hole mass from the Mbh-Mbulge relation.
% Fig. 1 of \citealt{Miller+15} plots the observed
% relationship of unresolved x-ray luminosity versus galactic stellar
% mass for a sample of quiescent galaxies. We may construct a similar
% relationship with our own sample using the inferred accretion rates
% from our solutions.  To infer an x-ray luminosity for each solution,
% we scale our calculated $\Mdot c^2$ by a constant. We pick $5\times
% 10^{-7}$, which gives x-ray luminosities comparable to
% those in \citealt{Miller+15}. We also try an alternate
% prescription in which $L_x$ is quadratic in $\Mdot$. In particular, we
% take $L_x=2 \times 10^{-4} (\eddr) \Mdot c^2$.

>>>>

This paper is organized as followed.  In $\S\ref{sec:model}$ we
describe our model, including the sample of galaxies used in our
analysis ($\S\ref{sec:gal_model}$) and our numerical procedure for
calculating the steady-state hydrodynamic profile of the CNM
($\S\ref{sec:hydro}$).  In $\S\ref{sec:results}$ we describe some
analytic results, while in $\S\ref{sec:numerical}$ we describe our
numerical results.  In $\S\ref{sec:discussion}$ we discuss the
implications of our results, to topics including...  In
$\S\ref{sec:conclusions}$ we summarize our conclusions.  Table
\ref{table:definitions} provides the definitions of commonly used
variables.  Appendices \ref{app:rs} and \ref{app:be} provides useful
analytic results, while Appendix \ref{app:windheat} describes our
method for calculating stellar wind heating and mass input in greater
detail.

@@

In the opposite limit of weak heating ($v_{w} \ll \sigma$), the
stagnation radius moves to large radii $\gg r_{\rm soi}$ or may not
exist at all, formally implying that the entire ISM of the galaxy is
accreting.  This result makes intuitive sense: gas is supplied to the
nuclear by stars which are gravitationally bound to the black hole, so
outflows are possible only if the specific heating rate $\sim
v_{w}^{2}$ significantly exceeds the specific gravitational binding
energy $\sim \sigma^{2}$.  In fact, for core galaxies the specific
heating rate necessary to produce an unbound outflow can greatly
exceed $\sigma^{2}$ due to the additional gravitational potential of
stars outside the sphere of influence (Appendix~\ref{app:be}).

@@

Thermal stability ($\dot{q}_{\rm heat} > |\dot{q}_{\rm rad}|$; $t_{\rm
  cool} > t_{\rm ff}$) near the stagnation radius thus requires a
critical minimum heating rate
\begin{align}
v_{w} > v_{\rm TI} \simeq
  {\rm max}\begin{cases}
   200 \eta_{0.02}^{0.14} M_{\bullet,8}^{0.11}\,{\rm km\,s^{-1}}  &, \text{core}\\
   200 \eta_{0.02}^{0.17} M_{\bullet,8}^{0.08}\,{\rm km\,s^{-1}}   &, \text{cusp}\\
f(r_b/r_{\rm soi})\sigma,     
  \end{cases}
  \label{eq:cooling3}
\end{align}
The requirement $v_{\rm w} \gtrsim \sigma$ results because the top two
lines are derived using equation (\ref{eq:rs_simple}) for $r_s$.
However, when $v_{\rm w} \lesssim f(r_b/r_{\rm soi})\sigma$ the
stagnation radius moves to very large radii, or may not exist at all
(see also Fig.~\ref{fig:stag}), resulting in much higher gas densities
and inevitable cooling instability.  The dimensionless function f is
approximately unity for cusp galaxies, but increases $\propto
(r_b/r_{\rm soi})^{1/2}$ for core galaxies (Fig.~\ref{fig:vwCrit}).  
%Our numerical results in $\S\ref{sec:numerical}$ demonstrate that $\tilde{v}_{w} \gtrsim v_{\rm CI} \sim 200-300$ km s$^{-1}$ is generally sufficient for stability at all radii (Fig.~\ref{fig:cooling}).

@@

If such cyclic outbursts also manifest on the properties of gas on
small, parsec scales, then the large scatter in the accretion rates
and hence X-ray luminosities in Fig \ref{fig:miller} could result from
the wide range of inflow rates experienced over the course of a cyclic
episode between instabilities and periodic nuclear outbursts
(e.g.~\citealt{Ciotti+10}).  Due to their deeper potential well, we
find that core galaxies are more likely to possess higher accretion
rates and potentially be thermally unstable for a fixed wind heating
rate than cusp galaxies (Fig.~XXX).  This could help explain why core
galaxies show a wide range of nuclear radio and X-ray luminosities
(suggesting cyclic accretion episodes), while cusp galaxies are
confined below a threshold (e.g., \citealt{Bender+89, Pellegrini99,
  Capetti&Balmaverde05}), consistent with thermally-stable,
steady-state picture. ({\bf AG: This picture may be dubious...Review
  the literature but I think that the higher upper bound for cores is
  just due to a mass selection effect.})

@@

% \subsubsection{Comparison with observed X-ray Luminosity-Mass
%   Relation}
% %Note currently I reverse-engineer the stellar mass from the black
% %hole mass inferred from M-sigma; however, it would be better to use
% %the tabulated black hole mass from the Mbh-Mbulge relation.
% Fig.1 of \citealt{Miller+15} plots the observed
% relationship of unresolved x-ray luminosity versus galactic stellar
% mass for a sample of quiescent galaxies. We may construct a similar
% relationship with our own sample using the inferred accretion rates
% from our solutions.  To infer an x-ray luminosity for each solution,
% we scale our calculated $\Mdot c^2$ by a constant. We pick $5\times
% 10^{-7}$, which gives x-ray luminosities comparable to
% those in \citealt{Miller+15}. We also try an alternate
% prescription in which $L_x$ is quadratic in $\Mdot$. In particular, we
% take $L_x=2 \times 10^{-4} (\eddr) \Mdot c^2$.
